\LARGE \bf Experimental Validation of Reference Spreading for Robotic Manipulation of Unmodeled Objects

### 2022-12-05 PaperPreliminary.tex
3-10

    with inertia matrix $M$, joint accelerations $\ddot{q}$, centrifugal, coriolis and gravity terms $h$, and commanded torque $\tau_{cmd}$. The robot has $n$ contacts, and for each contact i there is an external contact wrench $f_{c,i}$. The wrench's contribution on joint level is related through contact jacobian $J_i$ for which it holds that $\begin{bmatrix} \omega & \dot{p} \end{bmatrix}_i^T=J_i\dot{q}$, with $\omega$ the angular velocities of the contact body and $\dot{p}$ the Cartesian velocities of the contact point.

@@

    The desired impedance behavior is
    \begin{equation} \label{eq:impedance_desired}
     \Lambda  \begin{bmatrix} d_{\dot{\omega}} - \dot{\omega}   \\ d_{\ddot{p}} - \ddot{p}  \end{bmatrix} + D \begin{bmatrix}d_\omega - {\omega} \\ d_{\dot{p}} - \dot{p} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\d_p - {p}  \end{bmatrix} = d_{f_c} -f_c:
     \end{equation}

    a mass spring damper with with Cartesian-space inertia, damping and stiffness matrices $\Lambda$, $D$ and $K$, and rotation tracking error $e_{rot}$ as defined in Appendix xxx. Desired values are denotes as $d_{(\cdot)}$, e.g., the desired value for $\ddot{p}$ is  $d_{\ddot{p}}$.

@@

    Based on \ref{eq:impedance_desired} we can determine target task-space accelerations, $t_{\dot{\omega}}$ and $t_{\ddot{p}}$, following

    \begin{align} \begin{bmatrix}t_{\dot{\omega}}\\ t_{\ddot{p}}\end{bmatrix} &= \begin{bmatrix} {\dot{\omega}}\\{\ddot{p}}\end{bmatrix}+\Lambda^{-1} f_c\\
     &=     \begin{bmatrix} d_{\dot{\omega}}    \\ d_{\ddot{p}}  \end{bmatrix} +\Lambda^{-1} \left ( D \begin{bmatrix}d_\omega - {\omega} \\ d_{\dot{p}} - \dot{p} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\d_p - {p}  \end{bmatrix}- d_{f_c} \right ).
    \end{align}

@@

    \label{target_defs}
      \centering
    \begin{tabular}{ccc|cccc|}
    \cline{4-7}
                                      &                                    &               & \multicolumn{4}{c|}{Intermediate}                                              \\ \cline{2-7} 
    \multicolumn{1}{c|}{}             & \multicolumn{1}{c|}{Ante}          & Post          & 0             & 1         & 2             & 3                                  \\ \hline
    \multicolumn{1}{|c|}{$p_t$}       & \multicolumn{1}{c|}{$p^a_r$}       & $p^p_r$       & $p^p_r$       & $p^a_r$   & $p^a_r$       & $\gamma p^a_r+(1-\gamma)p^p_r$     \\
    \multicolumn{1}{|c|}{$\dot{p}_t$} & \multicolumn{1}{c|}{$\dot{p}^a_r$} & $\dot{p}^p_r$ & $\dot{p}^p_r$ & $\dot{p}$ & $0$ & $\gamma p + (1-\gamma)\dot{p}^p_r$ \\
    \multicolumn{1}{|c|}{$F_t$}       & \multicolumn{1}{c|}{$F^a_r$}       & $F^p_r$       & $F^p_r$       & $F^a_r$   & $F^a_r$       & $\gamma F^a_r + (1-\gamma)F^p_r$   \\ \hline
    \end{tabular}
    \end{table}

>>>>

    with inertia matrix $M$, joint accelerations $\ddot{q}$, centrifugal, coriolis and gravity terms $h$, and commanded torque $\tau_{cmd}$. The robot has $n$ contacts, and for each contact i there is an external contact wrench $f_{c,i}$. The wrench's contribution on joint level is related through contact jacobian $J_i$ for which it holds that $\begin{bmatrix} \omega & v \end{bmatrix}_i^T=J_i\dot{q}$, with $\omega$ the angular velocities of the contact body and $v$ the Cartesian velocities of the contact point $\dot{p}$.

@@

    The desired impedance behavior is
    \begin{equation} \label{eq:impedance_desired}
     \Lambda  \begin{bmatrix} \alpha_d - \alpha   \\ a_d - a  \end{bmatrix} + D \begin{bmatrix} \omega_d - {\omega} \\ v_d - v \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\p_d - {p}  \end{bmatrix} = {f_{c,d}} -f_c:
     \end{equation}

    a mass spring damper with with Cartesian-space inertia, damping and stiffness matrices $\Lambda$, $D$ and $K$, and rotation tracking error $e_{rot}$ as defined in Appendix xxx. Subscript $d$ denotes the desired value. Furthermore $\alpha$ and $a$ refer to the acceleration in rotation and translation respectively.

@@

    Based on \ref{eq:impedance_desired} we can determine target task-space accelerations, $\alpha_t$ and $a_t$, following

    % \begin{equation} 
$$
    \begin{bmatrix} \alpha_{t} \\ a_{t} \end{bmatrix} 
    = 
    \begin{bmatrix} \alpha \\ a \end{bmatrix} +
    \Lambda^{-1} f_{c}
$$
    $$
     =     \begin{bmatrix} \alpha_d    \\ a_d  \end{bmatrix} +\Lambda^{-1} \left ( D \begin{bmatrix} \omega_d - {\omega} \\ v_d - v \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\p_d - {p}  \end{bmatrix}- f_{c,d} \right ).
$$
    % \end{equation}

@@

    \label{target_defs}
      \centering
    \begin{tabular}{ccc|cccc|}
    \cline{4-7}
                                      &                                    &               & \multicolumn{4}{c|}{Intermediate}                                              \\ \cline{2-7} 
    \multicolumn{1}{c|}{}             & \multicolumn{1}{c|}{Ante}          & Post          & 0             & 1         & 2             & 3                                  \\ \hline
    \multicolumn{1}{|c|}{$p_t$}       & \multicolumn{1}{c|}{$p^a_r$}       & $p^p_r$       & $p^p_r$       & $p^a_r$   & $p^a_r$       & $\gamma p^a_r+(1-\gamma)p^p_r$     \\
    \multicolumn{1}{|c|}{$v_t$} & \multicolumn{1}{c|}{$v^a_r$} & $\dot{p}^p_r$ & $v^p_r$ & $v$ & $0$ & $\gamma v + (1-\gamma)v^p_r$ \\
    \multicolumn{1}{|c|}{$F_t$}       & \multicolumn{1}{c|}{$F^a_r$}       & $F^p_r$       & $F^p_r$       & $F^a_r$   & $F^a_r$       & $\gamma F^a_r + (1-\gamma)F^p_r$   \\ \hline
    \end{tabular}
    \end{table}

### 2022-12-05 Progress_13.tex
3-10

% Summary 
  \section*{Progress meeting 12 | 18 November, 2022}

  \section*{1. Progress}
  \begin{itemize}
  \item \textbf{Reference spreading with impedance control:}

  \begin{itemize}
      \item Reference spreading controllers:
        Last time I was having issues with getting the controller to work with a force reference. This was a programming issue which has since been fixed. In II.B of the paper attached at the end of this report, I give the general description of this controller. In IV.C, the control behaviour during ante, intermediate, and post impact phase are further specified. Section \ref{sec:results} shows the results. Here is are some things I found interesting about the results:
        \begin{itemize}
          \item Reference spreading reduces the control effort during non-nominal impacts
          \item Without a way to detect when contact is completed, introducing an intermediate impact phase with a predetermined length is not always beneficial.
          \item Disabling damping during the intermediate phase reduces control effort at the cost of worse tracking.
          \item A newly proposed intermediate control approach effectively blends the ante- and post impact reference for a balance between tracking and control effort.
        \end{itemize}
      \item Writing: So far I have written part of the introduction, an explanation of the impedance QP controller, and various options for the intermediate impact phase controller. The draft paper is attached at the end of this document.

@@

  \section*{2. Agenda}
  During the meeting, I would like to discuss the following points:
  \item Reference spreading results: There is still some room for improvement in the RS controllers. Jari already suggested trying more different intermediate modes and adding acceleration/force feedforward. I'm curious to hear if you have any other suggestions, and what you thinks should take priority in the limited time that remains. 
  \item What experiments should be featured in the final paper? I think the stamping experiment is the easiest to understand and therefore best at demonstrating the value of reference spreading. Box grabbing is good at demonstrating near-simultaneous impacts (i.e., first one arm makes contact, then the other). I think these two experiments capture all that is necesary regarding reference spreading, but they do not capture the full capability of the teleoperation framework.
      \item Planning: 
      \begin{itemize}
        \item We should schedule a date for the final defence. Must be ultimately January 22nd, unless I ask for extension. 
        \item (who should I reach out to for a possible deadline extension?)
        \item What is the final submission deadline for the thesis? Is two weeks in advance reasonable?
        \item Can I submit a draft version to receive your feedback? If so, we should schedule a deadline for this as well.
      \end{itemize}
  \end{itemize}

@@

  \begin{itemize}
      \item Finish finetuning of controller. One parameter that should be considered is the intermediate mode duration. Other improvements may be required depending on the discussion.
      \item Conduct final experiments. 
      \item Finish draft of paper for feedback. Some major components that should still be written: teleoperation, custom end effector, experimental results, conclusion.
      \item (Low priority for now: document controller code so that it may be used by others in the future)
  \end{itemize}

  \section*{4. Long-term planning}
  Shown below is the current long-term planning for the project phase.  The contact phase controller still needs some finetuning, after which I can finish data collection.

  The last phase of writing the report deponds on a few things:
  \begin{itemize}
    \item What will be the date of the final defence? (must be ultimately January 22nd)
    \item What is the deadline for my final thesis?
    \item Can I submit a draft of my thesis for you to review? What should be the deadline for this draft?
  \end{itemize}
  \begin{figure}[H]
  \centering
  \includegraphics[width=0.7\textwidth, trim={0.87cm 8.5cm 10cm 1.5cm},clip]{Graphics/planning v2.pdf}

@@

  \item[2b] \textbf{Contact monitoring:} When investigating contact monitoring, two approaches can be taken: either using proprioceptive or exteroceptive sensors. A possible improvement for contact monitoring using proprioceptive sensors could be to wait a fixed time starting from the last detected impact, rather than waiting a fixed time from the first impact. As for exteroceptive sensors, they can be a hurdle for large-scale commercial applications as they are not integrated in the robot. However, if a soft cover is to be mounted to the end effector, including tactile sensors for impact and contact monitoring becomes more feasible. Practical questions such as which tactile sensor to use and how to integrate it could be addressed, though this is not absolutely necessary for completing the research goals, and therefore has a low priority.
  \end{enumerate}
% main text
\newpage
  \section{Reference spreading results}\label{sec:results}
  The impedance controller (section II.B of appendix) was used with various reference spreading configurations (section IV.C of appendix). This setup was evaluated during a stamping task on the panda setup. Figure \ref{fig:stamp} shows the results.

  The command force $F_{cmd}$ will be used to evaluate the control effort. Without reference spreading, there is a clear peak in controll effort related to the peak in the velocity tracking error. The option RS inter0 does not suffer from this peak due to switching of the reference. None of the intermediate modes outperform RS inter0 for two reasons. Firstly, there is only one clearly visible impact while the intermediate mode is supposed to address near-simultaneous impacts. Furthermore, it is assumed that the intermediate mode lasts 0.2 seconds, whereas contact is completed sooner than that.

  Similarly, the controller was tested for a grabbing task using two arms. The results for the individual arms is shown in Figure \ref{fig:grab1} and \ref{fig:grab2}. In this scenario, the effect of the intermediate mode is more obvious since there are near-simultaneous impacts: the impact of arm 1, and the impact of arm 2. Intermediate options 1 and 2 experience a peak in the position tracking error which is based on the post-impact reference. This problem is resolved in option 3 thanks to the blending. Intermediate option 3 arguably has the best tradeoff between tracking performance and controll effort moderation.

\begin{figure}[]
  \centering
  \begin{adjustwidth}{0pt}{5pt}
  \centering
  \includegraphics[width=0.75\textwidth]{Graphics/Stamp result in z-direction_z_1.pdf}\\
    \includegraphics[width=0.75\textwidth]{Graphics/Stamp result in z-direction_z_2.pdf}
  \end{adjustwidth}
  \caption{Results of a stamping experiment. The top figure shows the results for the various approaches, both over the entire experiment, and over a small timeframe around the impact. The bottom figure isolates each approach for additional clarity. The dotted lines indicate the target as described in the control approaches.}
  \label{fig:stamp}
  \end{figure}

\begin{figure}[]
  \centering
  \begin{adjustwidth}{0pt}{5pt}
  \centering
  \includegraphics[width=0.75\textwidth]{Graphics/Boxgrab result in y-direction of arm 1_z_1.pdf}\\
    \includegraphics[width=0.75\textwidth]{Graphics/Boxgrab result in y-direction of arm 1_z_2.pdf}
  \end{adjustwidth}
  \caption{Results of arm 1 in boxgrab experiment. (no RS and RS inter0 show similar results. This might be a result of the nominal impact time being identical to the actual impact time, or an error during data processing. This requires further investigation.)}
  \label{fig:grab1}
  \end{figure}

\begin{figure}[]
  \centering
  \begin{adjustwidth}{0pt}{5pt}
  \centering
  \includegraphics[width=0.75\textwidth]{Graphics/Boxgrab result in y-direction of arm 2_z_1.pdf}\\
    \includegraphics[width=0.75\textwidth]{Graphics/Boxgrab result in y-direction of arm 2_z_2.pdf}
  \end{adjustwidth}
  \caption{Results of arm 2 in boxgrab experiment.(no RS and RS inter0 show similar results. This might be a result of the nominal impact time being identical to the actual impact time, or an error during data processing. This requires further investigation.)}
  \label{fig:grab2}
  \end{figure}

>>>>

% Summary 
  \section*{Progress meeting 13 | 6 December, 2022}

  \section*{1. Progress}
  \begin{itemize}
  \item \textbf{Controller finishing touches:}
      \begin{itemize}
        \item The controller now uses feedforward, which improves tracking prior to the impact. To achieve this, the impedance force during the demonstration is recorded and applied as feedforward during the replay.
        \item The impedance stiffness during replay was increased from 800 N/m to 1500 N/m. The stiffness during demonstration remains at 800 N/m. Combined with the feedforward, this causes the position tracking during the stamping task to decrease from approx. 20 mm to 3 mm prior to impact.
        \item Previously, reference spreading was only applied for the translational DoFs. I have now extended this so that reference spreading also works for the rotational DoFs.

@@

      \item \textbf{Data collection:} The stamping and grabbing experiment I showed in the previous progress report were repeated with the modified controller. Based on your feedback, this time I did stamping experiments where the impact surface was lowered, raised, or kept equal. Furthermore I recorded videos of the experiments.

      During the previous progress meeting, the results caused a bit of confusion; the surface was raised, yet the impact occured later due to poor tracking. Feedforward and higher gains solved this issue. The conclusions that can be drawn from the experiments did not change with respect to last time. The videos and figures are available at \url{https://surfdrive.surf.nl/files/index.php/s/3FMyF4ZadnimjsO}. In section \ref{sec:2}, two selected figures are shown: one for the stamping experiment where the surface was raised, and for box grabbing.

      \item \textbf{Writing:} Jari advised me to focus on determining the structure of my paper, rather than jumping straight into writing the individual sections. Section \ref{sec:1} shows what I have in mind. The paper in its current state is attached at the end of this document.

  \section*{2. Agenda}
  During the meeting, I would like to discuss the following points:

      \begin{itemize}
        \item In my opinion, the new experimental results demonstrate the value of reference spreading. Nevertheless, I think that reference spreading would be more effective with faster impact detection, which I think could be achieved by a momentum observer with different tuning from the tuning provided by franka emika. Do you think that I should spend time on a momentum observer, or spend more time on other changes to the controller? Or do you think that the results are already sufficiently convincing?
  \item I would like to go over the paper structure I have in mind. Jari has already given me some advice regarding what a paper should look like, however we came to the agreement that my final report should be somewhere between a paper and a thesis (to give an example, the custom end effector design wouldn't be considered a scientific contribution in an academic paper, whereas in a thesis report I believe it shouldn't be hidden in an appendix). I am curious to hear your opninion on this.
  \item I have made a planning for when I want to the send the report to you for feedback, and when I hope to receive the feedback. I want to ask if this planning seems reasonable to you.
  \end{itemize}

@@

  \begin{itemize}
      \item Make figures: error peaking, reference spreading, VR control loop with human, photo of setup, ...
      \item Document math: reference extending (also for rotations), impact detection, ...
      \item Finish draft of paper for feedback. Some major components that should still be written: teleoperation, custom end effector, experimental results, conclusion.
      \item (Low priority for now: document controller code so that it may be used by others in the future)
  \end{itemize}

  \section*{4. Long-term planning}
  Shown below is the current long-term planning for the project phase. It includes deadline, such as when I intend to submit the report for feedback. I received a deadline extension, meaning that the final deadline is the 5th of March. I plan for a defence around the 20th of February.

  \begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth, trim={0.87cm 6.5cm 4.5cm 1.5cm},clip]{Graphics/planning v2.pdf}

@@

  \item[2b] \textbf{Contact monitoring:} When investigating contact monitoring, two approaches can be taken: either using proprioceptive or exteroceptive sensors. A possible improvement for contact monitoring using proprioceptive sensors could be to wait a fixed time starting from the last detected impact, rather than waiting a fixed time from the first impact. As for exteroceptive sensors, they can be a hurdle for large-scale commercial applications as they are not integrated in the robot. However, if a soft cover is to be mounted to the end effector, including tactile sensors for impact and contact monitoring becomes more feasible. Practical questions such as which tactile sensor to use and how to integrate it could be addressed, though this is not absolutely necessary for completing the research goals, and therefore has a low priority.
  \end{enumerate}
% main text
\newpage
  \section{Paper structure}\label{sec:1}
  Below is the structure of the paper that I have in mind. The current stat of the paper is attached at the end of this document.

   Title: Experimental Validation of Reference Spreading for Robotic Manipulation of Unmodeled Objects
\begin{enumerate}
      \item Introduction

      \item System overview
\begin{enumerate}
      \item  Robot dynamics
      \item Task description (box weight/dimension, explanation of stamping and grabbing)
 \end{enumerate}
      \item Soft end effector design

      \item Trajectory planning using teleoperation
\begin{enumerate}
     \item  VR device

       \item  Impedance controller
\end{enumerate}
      \item Reference spreading
\begin{enumerate}
       \item  Impact detection

      \item  Reference formulation

       \item  Switching controller
\end{enumerate}
     \item  Experimental results
\begin{enumerate}
       \item  Stamping task

      \item  Grabbing task

       \item  (don’t think I should do more complicated tasks as they require more explanation while not showing additional benefits of RS)
      \end{enumerate}
     \item   Conclusion

     \item  Future work

\end{enumerate}
Appendix:

A.  End effector design: complete technical drawing and manufacturing process

B.  Experimental results: Can’t show all the data in the main text (probably only one or two direction, e.g. downward direction for stamping). Include full results in appendix.

C.  Math regarding rotation matrices and quaternions: definition of rotation error, interpolation of quaternions, integrating angular velocity for extending the rotation reference.

  \newpage

  \section{Selected figures of experimental results}\label{sec:2}
  (see following pages)
\begin{figure}[]
  \centering
  \begin{adjustwidth}{0pt}{5pt}
  \centering
  \includegraphics[width=0.75\textwidth]{Graphics/Stamp result in z-direction raised surface.pdf}\\
    \includegraphics[width=0.75\textwidth]{Graphics/Stamp result in z-direction raised surface_.pdf}
  \end{adjustwidth}
  \caption{Results of a stamping experiment. The top figure shows the results for the various approaches, both over the entire experiment, and over a small timeframe around the impact. The bottom figure isolates each approach for additional clarity. The dotted lines indicate the target as described in the control approaches.}
  \label{fig:stamp}
  \end{figure}

\begin{figure}[]
  \centering
  \begin{adjustwidth}{0pt}{5pt}
  \centering
  \includegraphics[width=0.75\textwidth]{Graphics/Grab result in y-direction_ Panda1.pdf}\\
    \includegraphics[width=0.75\textwidth]{Graphics/Grab result in y-direction_ Panda1_.pdf}
  \end{adjustwidth}
  \caption{Results of arm 1 in boxgrab experiment.}
  \label{fig:grab1}
  \end{figure}
